Symbolic Weighted Language Models and Parsing over Infinite Alphabets

### 2021-05-06 main.tex
details section VPA

\def\SWT{\textsf{swT}\xspace}
\def\SWA{\textsf{swA}\xspace}
\def\SWVPA{\textsf{sw-VPA}\xspace}
\def\weight{\mathsf{weight}}
\def\wei{\mathsf{w}}
\def\init{\mathsf{in}}
\def\final{\mathsf{out}}
\newcommand{\call}[1]{\ensuremath #1} %{\ensuremath \langle_{#1}}
\newcommand{\return}[1]{\ensuremath #1} %{\ensuremath {}_{#1}{\rangle}} % $\prescript{}{a}{)}$
\def\Sigmai{{\Sigma_\mathsf{i}}}
\def\Sigmac{{\Sigma_\mathsf{c}}}
\def\Sigmar{{\Sigma_\mathsf{r}}}
\def\Deltai{{\Delta_\mathsf{i}}}
\def\Deltac{{\Delta_\mathsf{c}}}
\def\Deltar{{\Delta_\mathsf{r}}}
\def\Phii{{\Phi_\mathsf{i}}}
\def\Phic{{\Phi_\mathsf{c}}}
\def\Phir{{\Phi_\mathsf{r}}}
\def\Phicr{{\Phi_\mathsf{cr}}}
\def\weii{{\wei_\mathsf{i}}}
\def\weic{{\wei_\mathsf{c}}}
\def\weir{{\wei_\mathsf{r}}}
\def\weie{{\wei_\mathsf{e}}}
\newcommand{\config}[2]{\ensuremath \genfrac{[}{]}{0pt}{}{#1}{#2}} 
\newcommand{\opluseq}{\ensuremath\mathrel{\oplus}=}
\newcommand{\ioi}[1]{\mathsf{ioi}({#1})}

@@

@@

\subsection{Label Theory}
\label{section:symbols}
 We shall now define the functions labeling the transitions of SW-automata and -transducers.
Let $\Sigma$ and $\Delta$ be two countable sets of symbols
called respectively input and output \emph{alphabets}, 
and let $\< \Semiring, \oplus, \zero, \otimes, \one>$ be a commutative semiring.

@@

\noindent
\marginpar{necessary?}
Finally, it is also required 
% that the codomain of every function of $\Phi_\Sigma$ and $\Phi_\Delta$ 
% is a subset of $\Phi_\epsilon$, and
that all partial applications of functions $\Phi_{\Sigma, \Delta}$, 
resp.  $f_a: y \mapsto f(a, y)$ for $a \in \Sigma$ and $y \in \Delta$
and  $f_b: x \mapsto f(x, b)$ for $b \in \Delta$ and $x \in \Sigma$, 
belong resp. to~$\Phi_\Sigma$ and~$\Phi_\Delta$.

@@

The symbolic-weighted transducer $T$ defines a mapping 
from the pairs of strings of $\Sigma^* \times \Delta^*$ 
into the weights of~$\Semiring$,
based on the following intermediate function $\weight_T$
defined recursively for every $q, q' \in Q$, 
%$a \in \Sigma$, $b \in \Delta$ 
for every strings of $s \in \Sigma^*$, $t \in \Delta^*$:
\[
\begin{array}{rccl}
\weight_T(q, s, t, q') & = & 
   \displaystyle\bigoplus_{q'' \in Q} &
   \wei(q, \epsilon, \epsilon, q'') \otimes \weight_T(q'', s, t, q')\\
   & \oplus & \displaystyle\bigoplus_{q'' \in Q} &
              \wei(q'', \epsilon, \epsilon, q') \otimes \weight_T(q, s, t, q'')\\
 & \oplus & \displaystyle\bigoplus_{\begin{array}{c}
                                      \scriptstyle q'' \in Q\\[-2pt]
                                      \scriptstyle s = au, a \in \Sigma
                                      \end{array}} &
   \wei(q, a, \epsilon, q'') \otimes \weight_T(q'', u, t, q')\\
 & \oplus & \displaystyle\bigoplus_{\begin{array}{c}
                                    \scriptstyle q'' \in Q\\[-2pt]
                                    \scriptstyle t = bv, b \in \Delta
                                    \end{array}} &
    \wei(q, \epsilon, b, q'') \otimes \weight_T(q'', s, v, q')\\ 
 & \oplus & \displaystyle\bigoplus_{\begin{array}{c}
                                    \scriptstyle q'' \in Q\\[-2pt]
                                    \scriptstyle s = au, a \in \Sigma\\[-2pt]
                                    \scriptstyle t = bv, b \in \Delta
                                    \end{array}} &
   \wei(q, a, b, q'') \otimes \weight_T(q'', u, v, q').\\ 
\end{array}
\]
We recall that, by convention, an empty sum with $\oplus$ is $\zero$. 
%
Since $\zero$ is absorbing for~$\otimes$ in~$\Semiring$,
one term $\wei(q, a, b, q'')$ equal to $\zero$ in the above expression 
will be ignored in the sum, meaning that there is no possible transition
from state $q$ into state $q'$ while reading $a$ and writing $b$ 
(including the case $a = \epsilon$ and $a = \epsilon$).
%These functions $\phi$ act as guards for the transducer's transitions, 
%preventing a transition when they return the absorbing $\zero$ of $\Semiring$.
This is analogous to a transition's guard not satisfied by $\<a, b>$ in 
the case of symbolic transducers.

The above expression of $\weight_T$ can be seen as a stateful definition of 
an edit-distance between a word $s \in \Sigma^*$ and a word $t \in \Delta^*$,
see also~\cite{Mohri03ijfcs}.
Intuitively, 
$\wei(q, a, \epsilon, q'')$, in the third line, is the cost of 
the deletion of a symbol $a \in \Sigma$ in $s$, 
$\wei(q, \epsilon, b, q'')$, in the fourth line, is the cost 
of the deletion of insertion of  $b \in \Delta$ in $s$, 
$\wei(q, a, b, q'')$, in the last line, is the cost 
of the substitution of  $a \in \Sigma$ by $b \in \Delta$.
The cost of a sequence of such operations transforming $s$ into $t$, 
is the product, with $\otimes$, of the individual costs of the operations involved;
And the distance between $s$ and $t$ is the sum, with $\oplus$,
of all such product of costs.

@@

\begin{example}
comparison of two sequences of timestamped events      
\end{example}

@@

\begin{proposition}
Given a \SWT $T$ 
over alphabet $\Sigma$, $\Delta$ and a commutative semiring~$\Semiring$, 
and $s \in \Sigma^*$, 
there exists an effectively constructible \SWA 
$A_{T, s}$ such that for all $t \in \Delta^*$, 
$A_{T, s}(t) = T(s, t)$.
\end{proposition}
%
\begin{proof}
Let $T = \< Q, \init, \vec{\wei}, \final >$ 
where $\vec{\wei} \< \wei_\epsilon, \wei_\Sigma, \wei_\Delta, \wei_{\Sigma, \Delta}>$
has an associated synthesized form $\wei$, 
and let $s = s_1 \ldots s_n$ with $s_1, \ldots, s_n \in \Sigma$.
%$n = |s|$ be the length of $s$.
%
Let $Q' = [1..n] \times Q$ be the state set of $A_{T, s}$,
and let, for all $q \in Q$, 
$\init'(1, q) = \init(q)$ and $\init'(i, q) = \zero$ for all $1 < i \leq n$, 
and $\final'(n, q) = \final(q)$ and $\final'(i, q) = \zero$ for all $1 \leq i <  n$.

\noindent
We define the functions $\wei'_\epsilon$ and $\wei'_\Sigma$ for $A_{T, s}$
in their synthesized form: for all $1 \leq i, j \leq n$, $q, q' \in Q$ and $a \in \Sigma$, 
and $b \in \Delta$, 
\[
\begin{array}{lclcl}
\wei'(\< i, q>, \epsilon, \< i, q'>)   & = & \wei(q, \epsilon, \epsilon, q'), & \quad & \\ 
\wei'(\< i, q>, \epsilon, \< i+1, q'>) & = & \wei(q, s_i, \epsilon, q') & & \mathrm{if~} i < n,\\ 
\wei'(\< i, q>, \epsilon, \< j, q'>)   & = & \zero & & \mathrm{if~} j \neq i, j \neq i+1,\\ 
\wei'(\< i, q>, b, \< i, q'>)          & = & \wei(q, \epsilon, b, q'),\\ 
\wei'(\< i, q>, b, \< i+1, q'>)        & = & \wei(q, s_i, b, q') & & \mathrm{if~} i < n,\\ 
\wei'(\< i, q>, b, \< j, q'>)          & = & \zero & & \mathrm{if~} j \neq i, j \neq i+1. 
\end{array}      
\]
Let $A_{T, s} = \< Q', \init', \vec{\wei'}, \final' >$.
\qed
\end{proof}

@@

@@

\subsection{Definition}
Let $\Sigma$ be a countable alphabet 
%finite (large) or infinite,
that we assume partitioned into three subsets 
$\Sigmai \uplus \Sigmac \uplus \Sigmac$ 
respectively called of \emph{internal}, \emph{call} and \emph{return}
symbols.
% \begin{itemize}
% \item a set $\Sigmai$ of \emph{internal symbols} denoted $a$,
% \item a set $\Sigmac$ of \emph{call symbols} denoted $\call{a}$,
% \item a set $\Sigmar$ of \emph{return symbols} denoted $\return{a}$.
% \end{itemize}
Let~$\< \Semiring, \oplus, \zero, \otimes, \one>$ be a commutative semiring and  
let  $(\Phi_\epsilon, \Phic, \Phir, \Phicr)$ be a label theory over $\Semiring$
%In order to simplify notations, %and following the definition of Section~\ref{section:transducer}, 
%we shall write respectively 
where $\Phic$, $\Phir$ and~$\Phicr$ stand respectively 
for~$\Phi_\Sigmac$, $\Phi_\Sigmar$ and~$\Phi_{\Sigmac, \Sigmar}$.
%
Moreover, we extend this theory with a set $\Phii$ 
of unary functions in $\Sigmai \to \Semiring$,
closed under $\oplus$ and $\otimes$.

\begin{definition}
A \emph{Symbolic Weighted Visibly Pushdown Automata} (\SWVPA)~$A$ 
over  $\Sigma = \Sigmai \uplus \Sigmac \uplus \Sigmar$ and $\Semiring$ is a tuple
$T = \< Q, P, \init, \weii, \weic, \weir, \weie, \final >$,
where $Q$ is a finite set of states, 
$P$ is a finite set of stack symbols, 
$\mathsf{in} : Q \to \Semiring$, 
respectively $\mathsf{out} : Q \to \Semiring,$
are functions defining the weight for entering, 
respectively leaving, a state, 
and 
$\weii : Q \times Q \to \Phii$,  
$\weic : Q \times Q \times P \to \Phic$,  
$\weir : Q \times P \times Q \to \Phicr$,  
$\weie : Q \times Q \to \Phir$,  
are transition functions.
\end{definition}
%
Similarly as in Section~\ref{section:transducer}, 
we extend the above transition functions as follows
for all $q, q' \in Q$, $p \in P$, 
$a \in \Sigmai$, 
$\call{c} \in \Sigmac$, 
$\return{r} \in \Sigmar$, 
overloading their names: % for simplicity:
\[
\begin{array}{lll}
\weii: Q \times \Sigmai \times Q \to \Semiring & 
\weii(q, a, q') = \phi_\mathsf{i}(a) & 
\mathrm{where~} \phi_\mathsf{i} = \weii(q, q'),\\
%
\weic: Q \times \Sigmac \times Q \times P \to \Semiring & 
\weic(q, \call{c}, q', p) = \phi_\mathsf{c}(\call{c}) & 
\mathrm{where~} \phi_\mathsf{c} = \weic(q, q', p),\\
%
\weir: Q \times \Sigmac \times P \times \Sigmar \times Q \to \Semiring & 
\weir(q, {\call{c}},  p, {\return{r}}, q') = \phi_\mathsf{r}({\call{c}},  {\return{r}}) & 
\mathrm{where~} \phi_\mathsf{r} = \weir(q, p, q'),\\
%
\weie: Q \times \Sigmar \times Q \to \Semiring & 
\weie(q, {\return{r}}, q') = \phi_\mathsf{e}({\return{r}}) &
\mathrm{where~} \phi_\mathsf{e} = \weie(q, q').\\
\end{array}      
\]

@@

We give now a formal definition of these transitions of the automaton $A$
in term of a weight value
computed by an intermediate function $\weight_A$.
In the case of a pushdown automaton, a configuration is composed 
of a state $q \in Q$ and a stack content $\theta \in \Theta^*$, where $\Theta = \Sigmac \times P$.
Therefore, $\weight_A$ is a function from 
$Q \times \Theta^* \times \Sigma^* \times Q \times \Theta^*$ into $\Semiring$.
\[
\begin{array}{rcl}
\weight_A\bigl(\config{q}{\theta}, a\, u, \config{q'}{\theta'}\bigr) & = & 
 {\displaystyle\bigoplus_{q'' \in Q}} \weii(q, a, q'') 
 \otimes \weight_A\bigl(\config{q''}{\theta}, u, \config{q'}{\theta'}\bigr)\\
%
\weight_A\bigl(\config{q}{\theta}, {\call{c}}\, u, \config{q'}{\theta'}\bigr) & = & 
 {\displaystyle\bigoplus_{\begin{array}{c}
                          \scriptstyle q'' \in Q\\[-2pt]
                          \scriptstyle p \in P
                          \end{array}}}
 \weic\bigl(q, {\call{c}}, q'', p\bigr) 
 \otimes \weight_A\bigl(\config{q''}{{\call{c}}\, p\cdot \theta}, u, \config{q'}{\theta'}\bigr)\\[1mm]
%
\weight_A\bigl(\config{q}{{\call{c}}\, p\cdot \theta}, {\return{r}}\, u, \config{q'}{\theta'}\bigr) & = & 
 {\displaystyle\bigoplus_{q'' \in Q}} 
 \weir\bigl(q, {\call{c}}, p, {\return{r}}, q''\bigr) 
 \otimes \weight_A\bigl(\config{q''}{\theta}, u, \config{q'}{\theta'}\bigr)\\
%
\weight_A\bigl(\config{q}{\bot}, {\return{r}}\, u, \config{q'}{\theta'}\bigr) & = & 
 {\displaystyle\bigoplus_{q'' \in Q}} \weie(q, {\return{r}}, q'') 
 \otimes \weight_A\bigl(\config{q''}{\bot}, u, \config{q'}{\theta'}\bigr)\\
\end{array}
\]
where $\bot$ denotes the empty stack and ${\call{c}}\, p\cdot \theta$ 
denotes a stack with the pair made of ${\call{c}}$ and $p$ on its top and $\theta$
as the rest of stack.

\noindent
The weight associated by $A$ to $s \in \Sigma^*$
is then defined as follows,
following empty stack semantics: 
\begin{equation}
A(s)  = 
{\displaystyle\bigoplus_{q, q' \in Q}} \textstyle
\mathsf{in}(q) \mathop{\otimes} 
\weight_A\bigl(\config{q}{\bot}, s, \config{q'}{\bot}\bigr) 
\mathop{\otimes} \mathsf{out}(q').
\label{eq:weightA}
\end{equation}

@@

\subsection{Properties}
The class of \SWVPA is closed under the binary operators of the underlying semiring.
%
\begin{proposition}
Let $A_1$ and $A_2$ be two (\SWVPA)
over the same $\Sigma$ and $\Semiring$.
There exists two $\SWVPA$ $A_1 \oplus A_2$ and $A_1 \otimes A_2$, 
effectively constructible, 
such that for all $s \in \Sigma^*$, 
$(A_1 \oplus A_2)(s) = A_1(s) \oplus A_2(s)$ and 
$(A_1 \otimes A_2)(s) = A_1(s) \otimes A_2(s)$.
\end{proposition}
The construction is essentially the same 
as the one of~\cite{dAntonyAlur14SVPDA}, in the case of the Boolean semiring.

\subsection{Best-first Search}
Assume a semiring $\Semiring$
commutative, complete, idempotent, superior, and total.
% total ? 
% monotonic and superior writ natural ordering
%Regarding the infinite sum operator, note that
%$\bigoplus_{x \in \Phi_\Sigma} \phi(x)$, 
%$\bigoplus_{y \in \Phi_\Delta} \psi(y)$, and  
%... exist and in $\Semiring$.
We propose a Dijkstra algorithm computing the minimal weight by $A$, 
\wrt $\leq_\oplus$, for a word in $\Semiring^*$.

>>>>

\def\SWT{\textsf{swT}\xspace}
\def\SWA{\textsf{swA}\xspace}
\def\SWVPA{\textsf{sw-VPA}\xspace}
\def\weight{\mathsf{weight}}
\def\wei{\mathsf{w}}
\def\init{\mathsf{in}}
\def\final{\mathsf{out}}
\newcommand{\call}[1]{\ensuremath #1} %{\ensuremath \langle_{#1}}
\newcommand{\return}[1]{\ensuremath #1} %{\ensuremath {}_{#1}{\rangle}} % $\prescript{}{a}{)}$
\def\Omegai{{\Omega_\mathsf{i}}}
\def\Omegac{{\Omega_\mathsf{c}}}
\def\Omegar{{\Omega_\mathsf{r}}}
\def\Sigmai{{\Sigma_\mathsf{i}}}
\def\Sigmac{{\Sigma_\mathsf{c}}}
\def\Sigmar{{\Sigma_\mathsf{r}}}
\def\Deltai{{\Delta_\mathsf{i}}}
\def\Deltac{{\Delta_\mathsf{c}}}
\def\Deltar{{\Delta_\mathsf{r}}}
\def\Phii{{\Phi_\mathsf{i}}}
\def\Phic{{\Phi_\mathsf{c}}}
\def\Phir{{\Phi_\mathsf{r}}}
\def\Phicr{{\Phi_\mathsf{cr}}}
\def\weii{{\wei_\mathsf{i}}}
\def\weic{{\wei_\mathsf{c}}}
\def\weir{{\wei_\mathsf{r}}}
\def\weie{{\wei_\mathsf{e}}}
\newcommand{\config}[2]{\ensuremath \genfrac{[}{]}{0pt}{}{#1}{#2}} 
\newcommand{\opluseq}{\ensuremath\mathrel{\oplus}=}
\newcommand{\ioi}[1]{\mathsf{ioi}({#1})}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% SWT & SWA
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\subsection{Label Theory}
\label{section:symbols}
We shall now define the functions labeling the transitions of SW-automata and -transducers.
Let $\Sigma$ and $\Delta$ be two countable sets of symbols
called respectively \emph{input} and \emph{output} \emph{alphabets}, 
and let $\< \Semiring, \oplus, \zero, \otimes, \one>$ be a commutative semiring.

@@

\noindent
%\marginpar{necessary?}
Finally, it is also required 
% that the codomain of every function of $\Phi_\Sigma$ and $\Phi_\Delta$ 
% is a subset of $\Phi_\epsilon$, and
that all partial applications of functions $\Phi_{\Sigma, \Delta}$, 
resp.  $f_a: y \mapsto f(a, y)$ for $a \in \Sigma$ and $y \in \Delta$
and  $f_b: x \mapsto f(x, b)$ for $b \in \Delta$ and $x \in \Sigma$, 
belong resp. to~$\Phi_\Sigma$ and~$\Phi_\Delta$.

@@

The symbolic-weighted transducer $T$ defines a mapping 
from the pairs of strings of $\Sigma^* \times \Delta^*$ 
into the weights of~$\Semiring$,
based on the following intermediate function $\weight_T$
defined recursively for every $q, q' \in Q$, 
%$a \in \Sigma$, $b \in \Delta$ 
for every strings of $s \in \Sigma^*$, $t \in \Delta^*$:
\[
\begin{array}{rccl}
\weight_T(q, s, t, q') & = & 
   \displaystyle\bigoplus_{q'' \in Q} &
   \wei(q, \epsilon, \epsilon, q'') \otimes \weight_T(q'', s, t, q')\\
   & \oplus & \displaystyle\bigoplus_{q'' \in Q} &
              \wei(q'', \epsilon, \epsilon, q') \otimes \weight_T(q, s, t, q'')\\
 & \oplus & \displaystyle\bigoplus_{\begin{array}{c}
                                      \scriptstyle q'' \in Q\\[-2pt]
                                      \scriptstyle s = au, a \in \Sigma
                                      \end{array}} &
   \wei(q, a, \epsilon, q'') \otimes \weight_T(q'', u, t, q')\\
 & \oplus & \displaystyle\bigoplus_{\begin{array}{c}
                                    \scriptstyle q'' \in Q\\[-2pt]
                                    \scriptstyle t = bv, b \in \Delta
                                    \end{array}} &
    \wei(q, \epsilon, b, q'') \otimes \weight_T(q'', s, v, q')\\ 
 & \oplus & \displaystyle\bigoplus_{\begin{array}{c}
                                    \scriptstyle q'' \in Q\\[-2pt]
                                    \scriptstyle s = au, a \in \Sigma\\[-2pt]
                                    \scriptstyle t = bv, b \in \Delta
                                    \end{array}} &
   \wei(q, a, b, q'') \otimes \weight_T(q'', u, v, q').\\ 
\end{array}
\]
%** The application of $\oplus$ can be seen as a ND choice between transitions **
We recall that, by convention, an empty sum with $\oplus$ is $\zero$. 
%
Since $\zero$ is absorbing for~$\otimes$ in~$\Semiring$,
one term $\wei(q, a, b, q'')$ equal to $\zero$ in the above expression 
will be ignored in the sum, meaning that there is no possible transition
from state $q$ into state $q'$ while reading $a$ and writing $b$ 
(including the case $a = \epsilon$ and $a = \epsilon$).
%These functions $\phi$ act as guards for the transducer's transitions, 
%preventing a transition when they return the absorbing $\zero$ of $\Semiring$.
This is analogous to a transition's guard not satisfied by $\<a, b>$ in 
the case of symbolic transducers.

The above expression of $\weight_T$ can be seen as a stateful definition of 
an edit-distance between a word $s \in \Sigma^*$ and a word $t \in \Delta^*$,
see also~\cite{Mohri03ijfcs}.
Intuitively, 
$\wei(q, a, \epsilon, q'')$, in the third line, is the cost of 
the deletion of a symbol~$a \in \Sigma$ in~$s$, 
$\wei(q, \epsilon, b, q'')$, in the fourth line, is the cost 
of the deletion of insertion of~$b \in \Delta$ in~$s$, 
and $\wei(q, a, b, q'')$, in the last line, is the cost 
of the substitution of  $a \in \Sigma$ by~$b \in \Delta$.
The cost of a sequence of such operations transforming $s$ into $t$, 
is the product, with $\otimes$, of the individual costs of the operations involved;
And the distance between $s$ and $t$ is the sum, with $\oplus$,
of all such product of costs.

@@

\begin{example}
DTW for sequences of timestamped events **     
\end{example}

\begin{example}
(simpler) poinwise distance between two sequences of timestamped events **
\end{example}

@@

\begin{proposition}
Given a \SWT $T$ 
over alphabet $\Sigma$, $\Delta$ and a commutative semiring~$\Semiring$, 
and $s \in \Sigma^*$, 
there exists an effectively constructible \SWA 
$A_{T, s}$ over $\Delta$ and $\Semiring$ 
such that for all $t \in \Delta^*$, $A_{T, s}(t) = T(s, t)$.
\end{proposition}
%
\begin{proof}
Let $T = \< Q, \init, \vec{\wei}, \final >$ 
where $\vec{\wei} \< \wei_\epsilon, \wei_\Sigma, \wei_\Delta, \wei_{\Sigma, \Delta}>$
has an associated synthesized form $\wei$, 
and let $s = s_1 \ldots s_n$ with $s_1, \ldots, s_n \in \Sigma$.
%$n = |s|$ be the length of $s$.
%
Let $Q' = [1..n] \times Q$ be the state set of $A_{T, s}$,
and let, for all $q \in Q$, 
$\init'(1, q) = \init(q)$ and $\init'(i, q) = \zero$ for all $1 < i \leq n$, 
and $\final'(n, q) = \final(q)$ and $\final'(i, q) = \zero$ for all $1 \leq i <  n$.

\noindent
We define the functions $\wei'_\epsilon$ and $\wei'_\Sigma$ for $A_{T, s}$
in their synthesized form: for all $1 \leq i, j \leq n$, $q, q' \in Q$ and $a \in \Sigma$, 
and $b \in \Delta$, 
\[
\begin{array}{lclcl}
\wei'(\< i, q>, \epsilon, \< i, q'>)   & = & \wei(q, \epsilon, \epsilon, q'), & \quad & \\ 
\wei'(\< i, q>, \epsilon, \< i+1, q'>) & = & \wei(q, s_i, \epsilon, q') & & \mathrm{if~} i < n,\\ 
\wei'(\< i, q>, \epsilon, \< j, q'>)   & = & \zero & & \mathrm{if~} j \neq i, j \neq i+1,\\ 
\wei'(\< i, q>, b, \< i, q'>)          & = & \wei(q, \epsilon, b, q'),\\ 
\wei'(\< i, q>, b, \< i+1, q'>)        & = & \wei(q, s_i, b, q') & & \mathrm{if~} i < n,\\ 
\wei'(\< i, q>, b, \< j, q'>)          & = & \zero & & \mathrm{if~} j \neq i, j \neq i+1. 
\end{array}      
\]
Note that the functions defined above belong to the set $\Phi_\Delta$ 
of the underlying label theory, 
by the hypothesis of closure under partial application.
%
The $\SWA$ wanted is $A_{T, s} = \< Q', \init', \vec{\wei'}, \final' >$.
\qed
\end{proof}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% SW-VPA
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\subsection{Definition}
Let $\Omega$ be a countable alphabet 
%finite (large) or infinite,
that we assume partitioned into three subsets 
$\Omegai$, $\Omegac$, $\Omegac$,
respectively called of \emph{internal}, \emph{call} and \emph{return}
symbols.
% \begin{itemize}
% \item a set $\Omegai$ of \emph{internal symbols} denoted $a$,
% \item a set $\Omegac$ of \emph{call symbols} denoted $\call{a}$,
% \item a set $\Omegar$ of \emph{return symbols} denoted $\return{a}$.
% \end{itemize}
Let~$\< \Semiring, \oplus, \zero, \otimes, \one>$ be a commutative semiring and  
let  $(\Phi_\epsilon, \Phic, \Phir, \Phicr)$ be a label theory over $\Semiring$
%In order to simplify notations, %and following the definition of Section~\ref{section:transducer}, 
%we shall write respectively 
where $\Phic$, $\Phir$ and~$\Phicr$ stand respectively 
for~$\Phi_\Omegac$, $\Phi_\Omegar$ and~$\Phi_{\Omegac, \Omegar}$.
%
Moreover, we extend this theory with a set $\Phii$ 
of unary functions in $\Omegai \to \Semiring$,
closed under $\oplus$ and $\otimes$.

\begin{definition}
A \emph{Symbolic Weighted Visibly Pushdown Automata} (\SWVPA)~$A$ 
over  $\Omega = \Omegai \uplus \Omegac \uplus \Omegar$ and $\Semiring$ is a tuple
$T = \< Q, P, \init, \weii, \weic, \weir, \weie, \final >$,
where $Q$ is a finite set of states, 
$P$ is a finite set of stack symbols, 
$\mathsf{in} : Q \to \Semiring$, 
respectively $\mathsf{out} : Q \to \Semiring,$
are functions defining the weight for entering, 
respectively leaving, a state, 
and 
$\weii : Q \times Q \to \Phii$,  
$\weic : Q \times Q \times P \to \Phic$,  
$\weir : Q \times P \times Q \to \Phicr$,  
$\weie : Q \times Q \to \Phir$,  
are transition functions.
\end{definition}
%
Similarly as in Section~\ref{section:transducer}, 
we extend the above transition functions as follows
for all $q, q' \in Q$, $p \in P$, 
$a \in \Omegai$, 
$\call{c} \in \Omegac$, 
$\return{r} \in \Omegar$, 
overloading their names: % for simplicity:
\[
\begin{array}{lll}
\weii: Q \times \Omegai \times Q \to \Semiring & 
\weii(q, a, q') = \phi_\mathsf{i}(a) & 
\mathrm{where~} \phi_\mathsf{i} = \weii(q, q'),\\
%
\weic: Q \times \Omegac \times Q \times P \to \Semiring & 
\weic(q, \call{c}, q', p) = \phi_\mathsf{c}(\call{c}) & 
\mathrm{where~} \phi_\mathsf{c} = \weic(q, q', p),\\
%
\weir: Q \times \Omegac \times P \times \Sigmar \times Q \to \Semiring & 
\weir(q, {\call{c}},  p, {\return{r}}, q') = \phi_\mathsf{r}({\call{c}},  {\return{r}}) & 
\mathrm{where~} \phi_\mathsf{r} = \weir(q, p, q'),\\
%
\weie: Q \times \Omegar \times Q \to \Semiring & 
\weie(q, {\return{r}}, q') = \phi_\mathsf{e}({\return{r}}) &
\mathrm{where~} \phi_\mathsf{e} = \weie(q, q').\\
\end{array}      
\]

@@

We give now a formal definition of these transitions of the automaton $A$
in term of a weight value
computed by an intermediate function $\weight_A$.
In the case of a pushdown automaton, a configuration is composed 
of a state $q \in Q$ and a stack content $\theta \in \Theta^*$, where $\Theta = \Omegac \times P$.
Therefore, $\weight_A$ is a function from 
$Q \times \Theta^* \times \Omega^* \times Q \times \Theta^*$ into $\Semiring$.
\[
\begin{array}{rcl}
\weight_A\bigl(\config{q}{\theta}, a\, u, \config{q'}{\theta'}\bigr) & = & 
 {\displaystyle\bigoplus_{q'' \in Q}} \weii(q, a, q'') 
 \otimes \weight_A\bigl(\config{q''}{\theta}, u, \config{q'}{\theta'}\bigr)\\
%
\weight_A\bigl(\config{q}{\theta}, {\call{c}}\, u, \config{q'}{\theta'}\bigr) & = & 
 {\displaystyle\bigoplus_{\begin{array}{c}
                          \scriptstyle q'' \in Q\\[-2pt]
                          \scriptstyle p \in P
                          \end{array}}}
 \weic\bigl(q, {\call{c}}, q'', p\bigr) 
 \otimes \weight_A\bigl(\config{q''}{{\call{c}}\, p\cdot \theta}, u, \config{q'}{\theta'}\bigr)\\[1mm]
%
\weight_A\bigl(\config{q}{{\call{c}}\, p\cdot \theta}, {\return{r}}\, u, \config{q'}{\theta'}\bigr) & = & 
 {\displaystyle\bigoplus_{q'' \in Q}} 
 \weir\bigl(q, {\call{c}}, p, {\return{r}}, q''\bigr) 
 \otimes \weight_A\bigl(\config{q''}{\theta}, u, \config{q'}{\theta'}\bigr)\\
%
\weight_A\bigl(\config{q}{\bot}, {\return{r}}\, u, \config{q'}{\theta'}\bigr) & = & 
 {\displaystyle\bigoplus_{q'' \in Q}} \weie(q, {\return{r}}, q'') 
 \otimes \weight_A\bigl(\config{q''}{\bot}, u, \config{q'}{\theta'}\bigr)\\
\end{array}
\]
where $\bot$ denotes the empty stack and ${\call{c}}\, p\cdot \theta$ 
denotes a stack where the pair made of ${\call{c}}$ and $p$ is the top symbol and $\theta$
is the rest of stack.

\noindent
The weight associated by $A$ to $s \in \Omega^*$
is then defined as follows,
following empty stack semantics: 
\begin{equation}
A(s)  = 
{\displaystyle\bigoplus_{q, q' \in Q}} \textstyle
\mathsf{in}(q) \mathop{\otimes} 
\weight_A\bigl(\config{q}{\bot}, s, \config{q'}{\bot}\bigr) 
\mathop{\otimes} \mathsf{out}(q').
\label{eq:weightA}
\end{equation}

@@

\subsection{Properties}
The class of \SWVPA is closed under the binary operators of the underlying semiring.
%
\begin{proposition}
Let $A_1$ and $A_2$ be two (\SWVPA)
over the same $\Omega$ and $\Semiring$.
There exists two $\SWVPA$ $A_1 \oplus A_2$ and $A_1 \otimes A_2$, 
effectively constructible, 
such that for all $s \in \Omega^*$, 
$(A_1 \oplus A_2)(s) = A_1(s) \oplus A_2(s)$ and 
$(A_1 \otimes A_2)(s) = A_1(s) \otimes A_2(s)$.
\end{proposition}
The construction is essentially the same 
as the one of~\cite{dAntonyAlur14SVPDA}, in the case of the Boolean semiring.

\subsection{Best-first Search}
Assume a semiring $\Semiring$
commutative, complete, idempotent, superior, and total.
% total ? 
% monotonic and superior writ natural ordering
%Regarding the infinite sum operator, note that
%$\bigoplus_{x \in \Phi_\Omega} \phi(x)$, 
%$\bigoplus_{y \in \Phi_\Delta} \psi(y)$, and  
%... exist and in $\Semiring$.
We propose a Dijkstra algorithm computing the minimal weight by $A$, 
\wrt $\leq_\oplus$, for a word in $\Semiring^*$.

\noindent
More precisely, 
let $\top$ be a fresh stack symbol which does not belong to $\Theta$,
and for every two states $q, q' \in Q$ and $\sigma \in \{ \bot, \top \}$, 
let 
\[ 
  d_0(q, \sigma, q') = \bigoplus_{s\in \Omega^*} 
  \textstyle
  \weight_A\bigl(\config{q}{\sigma}, s, \config{q'}{\sigma}\bigr). 
\]
Since $\Semiring$ is complete, this infinite sum is well defined, 
and since $\leq_\oplus$ is assumed total, it is the minimum in $\Omega^*$
of $s \mapsto \weight_A(\config{q}{\sigma}, s, \config{q'}{\sigma})$ \wrt this ordering.
%
When $\sigma = \bot$, $d_0(q, \sigma, q')$ is the central 
expression in a term of the definition (\ref{eq:weightA}) of $A(s_0)$ for the minimum $s_0$
(for the above function).
When $\sigma = \bot$, intuitively, it is the minimum weight of a computation of $A$
starting in state $q$ with a stack $\theta \in \Theta^*$ (possibly empty), 
and ending in state $q$ with the same stack $\theta$, such that moreover 
the computation does not touch a symbol of $\theta$. 
That means that during the computation, 
$A$ may apply the first case of in the definition of $\weight_A$ (internal symbol),
as well as the second case, to can push call symbols on the top of $\theta$, 
and may pop these symbols with the third case (return symbol). 
However, it cannot apply one of the two last cases (return symbol and empty stack)
when the current stack is $\theta$.
%pop symbols in $\theta$.
% Note that having a stack reduced to such a symbol makes impossible the application of the 
% two last cases in the definition of $\weight_A$ (return symbol and empty stack). 
% However, it is possible to apply the two first cases 
% (internal symbol or call symbol, with a push on the top of $\top$).
